### 2023-04-06 main.tex
ii

\section{Multi-species stirring process with reaction}
In \cite{casini2022uphill} it has been defined a process whose average occupation evolution equation are given by two Laplacians and a linear reaction term (mutation). This model consists in a sum of three Markov generators: a stirring, a stirring-mutation and a pure mutation one. For the lattice $\mathbb{Z}$, self-duality has been showed. In view of what have been proved in the previous section of this work, some questions arise naturally:
	\item Is it possible to extend this last duality result to the boundary-driven case in order to have an absorbing dual process?
	\item Does this process shares the same $gl(N)$ Lie-algebraic structure?
	\item Is it possible to extend the results to a case with arbitrary number of species $N\in \mathbb{N}$ and arbitrary maximal site occupation $\nu\in \mathbb{N}$?
The aim of this section is to give an answer to these questions. 
\subsection{Linear reaction-diffusion process with coupled discrete Laplacians}
Considering again a connected graph $G=(V,\mathcal{E})$ with set of vertices $V$ and set of edges $\mathcal{E}$, we introduce the generator
\begin{equation}\label{RDGenerator}
    \mathcal{L}^{rd}=\sum_{x,y\in\mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}^{rd}+\sum_{x\in V}\Gamma_{x}\mathcal{L}_{x}^{rd}
    \mathcal{L}_{x,y}^{rd}=\sigma_{11}\mathcal{L}_{x,y}+\sigma_{12}\sum_{c=1}^{N}\mathcal{L}_{x,y}^{c}+(\Upsilon-2\nu\sigma_{12})\mathcal{L}_{x,y}^{m}
with $\mathcal{L}_{x,y}$ is the generator the generator \eqref{edgeGenerator} and, for any function $f:\Omega\to \mathbb{R}$
    \mathcal{L}_{x,y}^{c}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{y}\left(f(\bm{n}-\delta_{a}^{x}+\delta_{g_{c}(b)}^{x}+\delta_{g_{c}(a)}^{y}-\delta_{b}^{y})-f(\bm{n})\right)
with the mapping 
    g_{c}(k)=\begin{cases}
    k+c \quad &\text{if}\quad k+c\leq N\\
    k+c-N\quad &\text{if}\quad k+c>N\\
    1\quad &\text{if}\quad k=1
\end{cases}\qquad \forall k\in \{1,\ldots N\}
    \mathcal{L}_{x,y}^{m}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}\left(f(\bm{n}-\delta_{a}^{x}+\delta_{b}^{x})-f(\bm{n})\right)
The site generator reads
	\mathcal{L}_{x}^{rd}f(\bm{n})=\sum_{a,b=1}^{N}\alpha_{a}^{x}n_{b}^{x}\left(f(\bm{n}-\delta_{b}^{x}+\delta_{a}^{x})-f(\bm{n})\right)
where the boundary parameters are defined as
	\alpha_{a}^{x}&=\sigma_{11}\nu\rho_{a}^{x}+\sigma_{12}\nu\sum_{b=2\,:\,b\neq a}^{N}\rho_{b}^{x}\qquad \text{if} \quad k\in \{2,\ldots,N\}\\
	\alpha_{1}^{x}&=\nu\left(\sigma_{11}+(N-1)\sigma_{12}\right)-\sum_{a=2}^{N}\alpha_{a}^{x}
\paragraph{Action on the occupation variable $n_{d}^{x}$ for the edge generator $\mathcal{L}_{x,y}^{rd}$}
The action on the occupation variable $n_{d}^{x}$ of particle type $d\in \{2,\ldots,N\}$ at site $x\in V $
is the following 
\begin{equation}\label{actionGraphRD}
    \mathcal{L}_{x,y}^{rd}n_{d}^{x}&=\nu \sigma_{11}(n_{d}^{y}-n_{d}^{x})+\nu\sigma_{12}\sum_{c=2\,:	,c\neq d}^{N}(n_{c}^{y}-n_{d}^{x})+(\Upsilon-2\nu\sigma_{12})\sum_{c=2}^{N}(n_{c}^{x}-n_{\alpha}^{x})

@@

	\alpha_{d}^{x}\left(\nu-n_{d}^{x}\right)-n_{d}^{x}\left(\nu(\sigma_{11}+(N-1)\sigma_{12})-\alpha_{d}^{x}\right)

>>>>

\section{Multi-species stirring process with reaction}
In \cite{casini2022uphill} it has been defined a process whose average occupation evolution equation are given by two Laplacians and a linear reaction term (mutation). This model consists in a sum of three Markov generators: a stirring, a stirring-mutation and a pure mutation one. For the lattice $\mathbb{Z}$, self-duality has been showed. In view of what have been proved in the previous section of this work, some questions arise naturally:
\begin{enumerate}
	\item Is it possible to extend this last duality result to the boundary-driven case in order to have an absorbing dual process?
	\item Does this process shares the same $gl(N)$ Lie-algebraic structure?
	\item Is it possible to extend the results to a case with arbitrary number of species $N\in \mathbb{N}$ and arbitrary maximal site occupation $\nu\in \mathbb{N}$?
\end{enumerate}
The aim of this section is to give an answer to these questions. 
\subsection{Linear reaction-diffusion process with coupled discrete Laplacians}
Considering again a connected graph $G=(V,\mathcal{E})$ with set of vertices $V$ and set of edges $\mathcal{E}$, we introduce the generator
\begin{equation}\label{RDGenerator}
    \mathcal{L}^{rd}=\sum_{x,y\in\mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}^{rd}+\sum_{x\in V}\Gamma_{x}\mathcal{L}_{x}^{rd}
\end{equation}
where
\begin{equation}
    \mathcal{L}_{x,y}^{rd}=\sigma_{11}\mathcal{L}_{x,y}+\sigma_{12}\sum_{c=1}^{N}\mathcal{L}_{x,y}^{c}+(\Upsilon-2\nu\sigma_{12})\mathcal{L}_{x,y}^{m}
\end{equation}
with $\mathcal{L}_{x,y}$ is the generator the generator \eqref{edgeGenerator} and, for any function $f:\Omega\to \mathbb{R}$
\begin{equation}
    \mathcal{L}_{x,y}^{c}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{y}\left(f(\bm{n}-\delta_{a}^{x}+\delta_{g_{c}(b)}^{x}+\delta_{g_{c}(a)}^{y}-\delta_{b}^{y})-f(\bm{n})\right)
\end{equation}
with the mapping 
\begin{equation}
    g_{c}(k)=\begin{cases}
    k+c \quad &\text{if}\quad k+c\leq N\\
    k+c-N\quad &\text{if}\quad k+c>N\\
    1\quad &\text{if}\quad k=1
\end{cases}\qquad \forall k\in \{1,\ldots N\}
\end{equation}
and where
\begin{equation}
    \mathcal{L}_{x,y}^{m}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}\left(f(\bm{n}-\delta_{a}^{x}+\delta_{b}^{x})-f(\bm{n})\right)
\end{equation}
The site generator reads
\begin{equation}
	\mathcal{L}_{x}^{rd}f(\bm{n})=\sum_{a,b=1}^{N}\alpha_{a}^{x}n_{b}^{x}\left(f(\bm{n}-\delta_{b}^{x}+\delta_{a}^{x})-f(\bm{n})\right)
\end{equation}
where the boundary parameters are defined as
\begin{equation}
	\begin{split}
	\alpha_{a}^{x}&=\sigma_{11}\rho_{a}^{x}+\sigma_{12}\sum_{b=2\,:\,b\neq a}^{N}\rho_{b}^{x}\qquad \text{if} \quad k\in \{2,\ldots,N\}\\
	\alpha_{1}^{x}&=\nu\left(\sigma_{11}+(N-2)\sigma_{12}\right)-\sum_{a=2}^{N}\alpha_{a}^{x}
	\end{split}
\end{equation}
and where, for every $x\in V$, we further assumed that
\begin{equation}
	\sum_{a=1}^{N}\rho_{a}^{x}=\nu
\end{equation}
\paragraph{Action on the occupation variable $n_{d}^{x}$ for the edge generator $\mathcal{L}_{x,y}^{rd}$}
The action on the occupation variable $n_{d}^{x}$ of particle type $d\in \{2,\ldots,N\}$ at site $x\in V $
is the following 
\begin{equation}\label{actionGraphRD}
\begin{split}
    \mathcal{L}_{x,y}^{rd}n_{d}^{x}&=\nu \sigma_{11}(n_{d}^{y}-n_{d}^{x})+\nu\sigma_{12}\sum_{c=2\,:	,c\neq d}^{N}(n_{c}^{y}-n_{d}^{x})+(\Upsilon-2\nu\sigma_{12})\sum_{c=2}^{N}(n_{c}^{x}-n_{\alpha}^{x})
\end{split}
\end{equation}

@@

	\alpha_{d}^{x}\left(\nu-n_{d}^{x}\right)-n_{d}^{x}\left(\nu(\sigma_{11}+(N-2)\sigma_{12})-\alpha_{d}^{x}\right)
	\\&=\alpha_{d}^{x}\nu-n_{d}^{x}\nu(\sigma_{11}+(N-2)\sigma_{12})
	\nu\sigma_{11}\rho_{d}^{x}+\sigma_{12}\left(\sum_{b=2\,:\,b\neq d}^{N}\rho_{b}^{x}\right)-n_{d}^{x}\nu(\sigma_{11}+(N-2)\sigma_{12})
